---
title: "Homework1"
author: "Jacqueline Oakes"
date: "2/1/2022"
output: github_document
---

# Problem 2.1
```{r setup, include = FALSE}
library(tidyverse)
library(tibble)

#Working directory for .RMD
knitr::opts_knit$set(echo = TRUE,
root.dir = rprojroot::find_rstudio_root_file())

#Controlling figure output in markdown
knitr::opts_chunk$set(fig.height = 6, 
                      fig.width = 6, 
                      fig.asp = .5, 
                      out.width = "90%",
                      out.height = "90%",
                      fig.align = "center",
                      cache = FALSE,
                      echo = TRUE)

#Set Theme for ggplot2
theme_set(theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom")) 

#Set Scientific notation output for knitr
options(scipen = 999)
```


```{r}
#create a tibble with mean and sd
normsample <- tibble(
            x = rnorm(1000, mean = 0, sd = 1),
            y = rnorm(1000, mean = 1, sd = 2))
normsample$sum_indicator <- ifelse(normsample$x + normsample$y > 0.5, TRUE, FALSE)

#use mutate to change T/F to Y/N
normsample <-  normsample %>%
   mutate(sum_indicator = as.factor(ifelse(sum_indicator == TRUE, "Yes", "No")))

#relevel so Y is before N 
normsample$sum_indicator <- forcats::fct_relevel(normsample$sum_indicator, "Yes")

#check work
view(normsample)

#create data visualization
normsample %>%
ggplot() +
  geom_point(aes(x = x, 
                 y = y, 
                 color = sum_indicator)) +
  scale_colour_discrete(name = "") +
  labs(x = "Sample1", y = "Sample2",
       title = "Independent Bivariate Normal Random Sample") +
    theme(plot.title = element_text(face = "bold")) +
    theme(panel.background = element_blank()) +
    theme(axis.line = element_line(colour = "black")) +
    theme(plot.title = element_text(hjust = 0.5))
```

# Problem 2.2
```{r, results='hide'}
penguin.df <- read_rds("./data/penguin.RDS")

str(penguin.df)
#see how many rows are in the dataset
nrow(penguin.df)
#see how many columns are in the dataset
ncol(penguin.df)

mean(penguin.df$bill_length_mm, na.rm = TRUE)
sd(penguin.df$bill_length_mm, na.rm = TRUE)

mean(penguin.df$flipper_length_mm, na.rm = TRUE)
sd(penguin.df$flipper_length_mm, na.rm = TRUE)

ggplot(na.omit(penguin.df)) +
  geom_point(aes(x = bill_length_mm, 
                 y = flipper_length_mm, 
                 color = species)) +
  labs(x = "Bill Length", y = "Flipper Length",
       title = "Penguins of Antarctica") +
    theme(plot.title = element_text(face = "bold")) +
    theme(panel.background = element_blank()) +
    theme(axis.line = element_line(colour = "black")) +
    theme(plot.title = element_text(hjust = 0.5))

ggplot(na.omit(penguin.df)) +
  geom_point(aes(x = bill_length_mm, 
                 y = flipper_length_mm, 
                 color = species)) +
  facet_grid(~sex, scales="free") +
  labs(x = "Bill Length", y = "Flipper Length",
       title = "Penguins of Antarctica") +
    theme(plot.title = element_text(face = "bold")) +
    theme(panel.background = element_blank()) +
    theme(axis.line = element_line(colour = "black")) +
    theme(plot.title = element_text(hjust = 0.5))
```

* The Penguin dataset for each observation shows the size measurements for adult foraging penguins near Palmer Station Antarctica.
* The number of rows is `r nrow(penguin.df)`
* The number of columns is `r ncol(penguin.df)`
* Each variable is a penguin located in Antarctica and the features of that specific penguin, such as, bill length and depth, flipper length, body mass, and sex.
* The mean penguin bill length is `r mean(penguin.df$bill_length_mm, na.rm = TRUE)`
* The standard deviation penguin bill length is `r sd(penguin.df$bill_length_mm, na.rm = TRUE)`
* The mean penguin flipper length is `r mean(penguin.df$flipper_length_mm, na.rm = TRUE)`
* The standard deviation penguin flipper length is `r sd(penguin.df$flipper_length_mm, na.rm = TRUE)`